The Leaky Semicolon

### 2021-10-27 paper/mca2.tex
Updates
removing todos
messing in 4.2

Lowering \PwTmcaTITLE{1} to \armeight{} requires a full fence after every
acquiring read.  To see why, consider the following attempted
execution, where the final values of both $x$ and $y$ are $2$.
\begin{gather*}
  \PW{x}{2}\SEMI 
  \PR[\mACQ]{x}{r}\SEMI
  \PW{y}{r{-}1} \PAR
  \PW{y}{2}\SEMI
  \PW[\mREL]{x}{1}
  \\
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{2}}{}
      \raevent{b}{\DR[\mACQ]{x}{2}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \event{d}{\DW{y}{2}}{right=2.5em of c}
      \raevent{e}{\DW[\mREL]{x}{1}}{right=of d}
      \rfint{a}{b}
      \sync{b}{c}
      \wk{c}{d}
      \sync{d}{e}
      \wk[out=-165,in=-15]{e}{a}
    \end{tikzinline}}
\end{gather*}
The execution is  allowed by \armeight, but disallowed by \PwTmca{1}, due to
the cycle.

@@

  \begin{enumerate}[topsep=0pt]
  \item[{\labeltext[\textsc{p}6a]{(\textsc{p}6a)}{par-le-rf1}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\bEv\xrfx\aEv$ then $\bEv\lt\aEv$,
  \item[{\labeltext[\textsc{p}6b]{(\textsc{p}6b)}{par-le-rf2}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\aEv\xrfx\bEv$ then $\aEv\lt\bEv$,
  \item[{\labeltext[\textsc{s}6a]{(\textsc{s}6a)}{seq-le-delays-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rdelays \labeling_2(\aEv)$ 
    then either $\bEv\xrfx\aEv$ or $\bEv\le\aEv$,
  \item[{\labeltext[\textsc{s}6b]{(\textsc{s}6b)}{seq-le-rf-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rblocks \labeling_2(\aEv)$
    then $\cEv\xrfx\aEv$ implies $\bEv\le\cEv$.
    % \\
    % if $\labeling_1(\cEv) \rblocks \labeling_2(\aEv)$ and $\bEv\xrfx\aEv$
    % then $\cEv\le\bEv$.
  % \item[{\labeltext[\textsc{i}6a]{(\textsc{i}6a)}{if-le-rf}}]
  %   ${\rrfx}\subseteq {\rrfx_1}\cup {\rrfx_2}$
  \end{enumerate}
  % The definition of \PwTmca{2} is a \PwT{} (\refdef{def:pomset}) equipped with an injective
  % relation $\rrfx$ that satisfies requirements \ref{pom-rf-match} and
  % \ref{pom-rf-block} of \refdef{def:pwt:mca1}.

@@

  %\input{fig-mca2.tex}
\end{definition}
% We write $\semmca{2}{}$ for the semantic function.
A \PwTmca{2} need not satisfy requirement \ref{pom-rf-le}, and thus we may
have $\bEv\xrfx\aEv$ and $\aEv\lt\bEv$.

@@

\todo{Example using \ref{seq-le-delays-rf} and \ref{seq-le-rf-rf}. Perhaps
  move \reflem{lem:mca2} to appendix.}

>>>>

Lowering \PwTmca{1} to \armeight{} requires a full fence after every
acquiring read.  To see why, consider the following attempted
execution, where the final values of both $x$ and $y$ are $2$.
\begin{gather*}
  \taglabel{internal-acq}
  \begin{gathered}
  \PW{x}{2}\SEMI 
  \PR[\mACQ]{x}{r}\SEMI
  \PW{y}{r{-}1} \PAR
  \PW{y}{2}\SEMI
  \PW[\mREL]{x}{1}
  \\
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{2}}{}
      \raevent{b}{\DR[\mACQ]{x}{2}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \event{d}{\DW{y}{2}}{right=2.5em of c}
      \raevent{e}{\DW[\mREL]{x}{1}}{right=of d}
      \rfint{a}{b}
      \sync{b}{c}
      \wk{c}{d}
      \sync{d}{e}
      \wk[out=-170,in=-10]{e}{a}
    \end{tikzinline}}
\end{gathered}
\end{gather*}
The execution is  allowed by \armeight, but disallowed by \PwTmca{1}, due to
the cycle.

@@

  \begin{enumerate}[topsep=0pt]
  \item[{\labeltext[\textsc{p}6a]{(\textsc{p}6a)}{par-le-rf1}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\bEv\xrfx\aEv$ then $\bEv\lt\aEv$,
  \item[{\labeltext[\textsc{p}6b]{(\textsc{p}6b)}{par-le-rf2}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\aEv\xrfx\bEv$ then $\aEv\lt\bEv$,
  \item[{\labeltext[\textsc{s}6a$'$]{(\textsc{s}6a$'$)}{seq-le-delays-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rdelays \labeling_2(\aEv)$ 
    then either $\bEv\xrfx\aEv$ or $\bEv\le\aEv$,
  \item[{\labeltext[\textsc{s}6b]{(\textsc{s}6b)}{seq-le-rf-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rblocks \labeling_2(\aEv)$
    then $\cEv\xrfx\aEv$ implies $\bEv\le\cEv$.
    % \\
    % if $\labeling_1(\cEv) \rblocks \labeling_2(\aEv)$ and $\bEv\xrfx\aEv$
    % then $\cEv\le\bEv$.
  % \item[{\labeltext[\textsc{i}6a]{(\textsc{i}6a)}{if-le-rf}}]
  %   ${\rrfx}\subseteq {\rrfx_1}\cup {\rrfx_2}$
  \end{enumerate}
  % The definition of \PwTmca{2} is a \PwT{} (\refdef{def:pomset}) equipped with an injective
  % relation $\rrfx$ that satisfies requirements \ref{pom-rf-match} and
  % \ref{pom-rf-block} of \refdef{def:pwt:mca1}.

@@

  %\input{fig-mca2.tex}
\end{definition}
\ref{par-le-rf1} and \ref{par-le-rf2} ensure that $\bEv\xrfx\aEv$ implies
$\bEv\lt\aEv$ when the actions come from different threads.  However, we may
have $\bEv\xrfx\aEv$ and $\aEv\lt\bEv$ within a thread, as between
$\DWP{x}{2}$ to $\DRP[\mACQ]{x}{2}$ in \ref{internal-acq}.  To allow this,
\ref{seq-le-delays-rf} relaxes \ref{seq-le-delays} from \PwTmca{1}.
\ref{seq-le-rf-rf} is then necessary to stop stale reads within a thread.
For example, it prevents a read of $1$ in
\begin{math}
  \PW{x}{1}
  \SEMI
  \PW{x}{2}
  \SEMI
  \PR{x}{r}.
\end{math}
\todo{Is s6b necessary?}

@@
